Technical Report

### 2018-05-22 fully-abstract-comp/technical-report.tex
Prove a compatibility theorem for linking, fix an odd cyclicity in the definitions with an appropriate later in the definition.

@@

\begin{align*}
  \lrcomp(W) &=
  \left\{\begin{multlined}
     \npair{\var{comp},\var{comp}} \;\mid \;\\
      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
      W \vdash \var{comp} \tand\\
      \text{For all } W' \future W \ldotp \text{If } \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'})\\
      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{multlined}
    \right\}\\
  &\cup \left\{
    \begin{multlined}
     \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
     \{(c_{\mathrm{main},c}, c_{\mathrm{main},c}),(c_{\mathrm{main},d}, c_{\mathrm{main},d})\} \subseteq (\overline{c_{\mathrm{export}}},\overline{c_{\mathrm{export}}})
       \end{multlined}
    \right\} 
\end{align*}

\begin{lemma}[Compatibility lemma for linking]
  If
  \begin{itemize}
  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$
  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$
  \end{itemize}
  then
  \begin{itemize}
  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1\oplus W_2)$.
  \end{itemize}
\end{lemma}

>>>>

\begin{lemma}[World specifies component pure]
  \label{lem:world-specifies-comp-pure}
  If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
\end{lemma}
\begin{proof}
  Follows easily from the definitions.
\end{proof}

@@

\begin{align*}
  \lrcomp(W) &=
  \left\{\begin{multlined}
     \npair{\var{comp},\var{comp}} \;\mid \;\\
      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
      W \vdash \var{comp} \tand\\
      \text{For all } W' \future W \ldotp \text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
  \end{multlined}
    \right\}\\
  &\cup \left\{
    \begin{multlined}
     \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
     \{(c_{\mathrm{main},c}, c_{\mathrm{main},c}),(c_{\mathrm{main},d}, c_{\mathrm{main},d})\} \subseteq (\overline{c_{\mathrm{export}}},\overline{c_{\mathrm{export}}})
       \end{multlined}
    \right\} 
\end{align*}

\begin{lemma}[Compatibility lemma for linking]
  If
  \begin{itemize}
  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$
  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$
  \end{itemize}
  then
  \begin{itemize}
  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W)$.
  \end{itemize}
\end{lemma}
\begin{proof}
  First consider the case where one of the two pairs of components have a pair of ``main'' capabilities.
  By the definition of $\lrcomp$, these capabilities have to be part of that components' export list.
  It is then easy to see that the other component cannot have ``main'' capabilities (otherwise linking is not defined) and it is sufficient to prove the result for the underlying components (without the ``main'' capabilities).
  Therefore, we can restrict ourselves to the case where both pairs of components are of the form $\var{comp}_0$, i.e. no ``main'' capabilities.

  By Lemma~\ref{lem:world-specifies-linked}, we have that $W \vdash \var{comp}_1\bowtie \var{comp}_2$.

  By definition of $\bowtie$, we have that the following hold:
  \begin{align*}
    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] &
    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
  \end{align*}

  Now take $W' \future W$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
  Then it remains to show that $\overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$.
  We do this by complete induction on $n$, so that we can assume that $\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.

  For each $$\npair{(c_{\mathrm{export}},c_{\mathrm{export}})} \in \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}}$$, we have that $\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}$ is either in $\overline{\npair{(c_{\mathrm{export},1},c_{\mathrm{export},1})}}$ or $\overline{\npair{(c_{\mathrm{export},2},c_{\mathrm{export},2})}}$.
  By unfolding $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$, it suffices to prove that $$\overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})$$ and $$\overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})$$
  for all $n' < n$.

  Simple set algebra tells us that 
  $$\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}$$
  and for all $n' < n$, we know that $$\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$$
  and
  $$\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}$$
\end{proof}
